\center \fontsize{18}{20} \selectfont Revealing the Landscape of Globally Color-Dual Multi-loop Integrands

### 2023-09-28 notes/2loopNLSM.tex
Thus -> Therefore
Prepping for arxiv upload.

\usepackage{tikz-feynman}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Comment the following out to compile new images
%
%\usetikzlibrary{external}  
%\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
%\tikzexternalize[                     %% Activate externalization
%prefix=images/,                    %% Avoid cluttering the directory
%system call={                       %% Use lualatex in system call
%  % pdflatex \tikzexternalcheckshellescape -halt-on-error -shell-escape -interaction=batchmode -jobname="\image" %"\texsource" || rm "\image.pdf"
%},
%]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\newcommand{\jm}[1]{\textcolor{blue}{\textbf{JM: }{#1}}}
\newcommand{\nhp}[1]{\textcolor{nhpRed}{\textbf{NHP: }{#1}}}

@@

Despite the tremendous success in leveraging this duality to
compute gauge and gravity observables to high orders in perturbation
theory \cite{FiveLoopN4, GeneralizedDoubleCopyFiveLoops, Bern:2018jmv,
  Carrasco:2021otn}, color-kinematics remains a conjecture at loop
level. Indeed, there are many examples in the literature where
identifying color-dual representations beyond one-loop has posed a
formidable challenge
\cite{Mogull:2015adi, Johansson:2017bfl, KalinN2TwoLoop, Bern:2015ooa}. In this work, we use
the nonlinear sigma model (NLSM) and Yang-Mills, two theories proven
to permit color-dual representations at tree-level
\cite{Feng:2010my,Cachazo:2014xea}, as case studies to advance our
understanding of the kinematic algebra at the multi-loop level.

We begin with an overview of color-kinematics duality in
\cref{background}, and define the notion of ``globally'' color-dual integrands in \cref{sec:bootstrap}. 
We then construct a manifestly color-dual theory in
\cref{sec:cubic}, which generates $D$-dimensional color-dual $n$-point
numerators at both tree-level and one-loop. When plugging in appropriate on-shell states, 
we find that these numerators underpin color-dual representations of self-dual Yang-Mills, NLSM and Chern-Simons theory
through one-loop. \jm{Mention CS here?} \nhp{Added sentence} The construction of this theory, which we dub
\textit{semi-abelian Yang-Mills}, relies on isolating the cubic sector
of pure Yang-Mills amplitudes from the four-point contacts that are
needed to fully realize non-abelian gauge symmetry. Despite the
potency of this theory through one-loop, the construction runs into an
obstruction at two-loop, which we describe in \cref{2LoopObstruction}.

@@

The color factor
$C_\Gamma$ is the string of structure constants $f^{abc}$ associated
with the graph $\Gamma$,\footnote{This paper is focused on theories
  with particles that transform in the adjoint.  However, the double
  copy can be applied to a far more diverse set of theories with
  particles that transform in the fundamental or more diverse
  representations, see for example \cite{Johansson:2017srf,
    Johansson2014zca, Johansson:2015oia, Johansson:2019dnu, Carrasco:2020ywq,Carrasco:2022jxn,
    Carrasco:2023vjg}.}  the denominator $d_\Gamma$
is the product of propagators associated with the graph $\Gamma$, and
the numerator $N_\Gamma$ is all of the remaining kinematic factors
including dot products of momenta and polarization vectors. In this 
form the color factors $C_\Gamma$ are not independent and are related to
one another via a set of Jacobi identities that look schematically
like
\begin{equation}
C_i+C_j+C_k=0,
\end{equation}
where $i$, $j$, and $k$ are three graphs.  Example Jacobi relations
are presented in detail in \cref{sec:bootstrap}. The duality between color and kinematics is the statement that that there exists a way to rearrange factors between
the various kinematic numerators $N_\Gamma$ such that the numerators
obey the same Jacobi identities as the color factors
\begin{equation}
N_i +N_j +N_k=0.
\end{equation}
If color-kinematics duality can be achieved at the off-shell level,
then the kinematic numerators should satisfy an algebra just like the
color factors.  In theories where the kinematic algebra is known, it
is often the algebra of volume preserving diffeomorphisms
\cite{Monteiro2011pc, Ben-Shahar:2021zww, Cheung:2020djz,
  Cheung:2021zvb, Cheung:2022mix}.
When the kinematic algebra is not known explicitly, the kinematic numerators are typically constructed from an ansatz.
As an example, the result of such a calculation for four-point NLSM scattering at tree level produces the numerators
\begin{align}\label{pionNums}
&N^{\text{NLSM}}_s = s(t-u)\\
&N^{\text{NLSM}}_t = t(u-s)\\
&N^{\text{NLSM}}_u = u(s-t)
\end{align}
where the coupling has been normalized away.  These numerators sum to
zero by explicit calculation.  Note that the $s$-channel numerator
contains an explicit factor of $s$ that cancels with the propagator
and the same goes for the $t$- and $u$-channels.
The net result is that the amplitude is a local function, just as one would expect for the pion tree
amplitude.

Once color-dual numerators $N_\Gamma$ have been found, they can be
used to replace the color factors ($C_\Gamma \to N_\Gamma$) in a
\emph{different} amplitude to produce the corresponding amplitude in
a new theory,
\be\label{DCconstruction}
\tilde{\mathcal{A}}_n^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int
\frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{C_\Gamma
  \tilde{N}_\Gamma}{d_\Gamma} ~~ \xrightarrow{(C_\Gamma \to N_\Gamma)}
~~ \mathcal{M}_n^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int
\frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{N_\Gamma
  \tilde{N}_\Gamma}{d_\Gamma}.
\end{equation}
This is known as the double-copy construction \cite{BCJ,Bern:2010ue,BCJreview}.
Importantly, the $\tilde{N}_\Gamma$ do not need to respect
color-kinematics duality.  The prototypical example of the double-copy
is that YM double copied with YM results in gravity where the gauge
invariance of each separate gluon produces the diffeomorphism
invariance of the graviton.  The double copy can be proven at tree
level using the Kawai-Lewellen-Tye (KLT) relations but there is significant evidence that
the double copy persists to all loop orders \cite{FiveLoopN4,
  GeneralizedDoubleCopyFiveLoops, Bern:2018jmv, Carrasco:2021otn, KLT,
  KiermaierTalk}.  In addition to pointing at undiscovered
structure hidden in the Lagrangian for gravity, the double copy is
immensely useful at a utilitarian level since it is much more
efficient at producing $\mathcal{M}_n^L$ than traditional methods like
Feynman rules.

Many of the one-loop results presented in \cref{sec:cubic} are informed
by the tree-level double-copy so we will review the tree case here.
The color structure of tree-level amplitudes is particularly simple.
A fully color dressed tree amplitude $\mathcal{A}_n$ can be decomposed
in terms of the trace basis
\begin{equation}
  \mathcal{A}_n = \sum \limits_{\sigma\in S_{n-1}} \text{Tr}(T^{a_{\sigma_1}} T^{a_{\sigma_2}}...T^{a_{\sigma_{n-1}}} T^{a_n}) A_n[\sigma_1, \sigma_2,...\sigma_{n-1}, n],
  \label{eq:trace-basis}
\end{equation}
where $T^a$ are generators of the gauge group and, because of the
cyclicity of the trace, one of the legs is held fixed so that the sum
runs over the $(n-1)!$ permutations of the remaining external legs.
The coefficients of the color factors are the color-ordered partial
amplitudes $A_n[...]$.  Any tree-level color factor can be converted
into a linear combination of Del Duca-Dixon-Maltoni (DDM) half-ladder
color factors
\begin{equation}
\text{DDM}[a_1, a_2,... a_n] \equiv f^{a_1 a_2 b_1} f^{b_1 a_3 b_2} f^{b_2 a_4 b_3}...f^{b_{n-3} a_{n-1} a_n}
\end{equation}
associated with the graph
\begin{equation}
\halfLadder
\end{equation}
by repeated application of the Jacobi identity \cite{DixonMaltoni}.
The fully color-dressed tree amplitude can then be re-expressed as
\begin{equation}
  \mathcal{A}_n = \sum \limits_{\sigma\in S_{n-2}} \text{DDM}[a_1, a_{\sigma_2}, a_{\sigma_3}, a_{\sigma_{n-1}}, a_n]  A_n[1, \sigma_2,...\sigma_{n-1}, n]
  \label{eq:ddm-basis}
\end{equation}
with legs $1$ and $n$ fixed.  Since the sum only contains $(n-2)!$
terms, there must be additional relations -- the
Kleiss-Kuijf (KK) relations \cite{Kleiss:1988ne} -- amongst the
partial amplitudes appearing in the right-hand side of
\cref{eq:trace-basis}.  The KK relations are generic to any theory
with purely adjoint particles.  Color-kinematics duality further
implies that the KK basis is overcomplete and can be reduced to a
basis of $(n-3)!$ amplitudes via the fundamental Bern-Carrasco-Johansson (BCJ) identities
\cite{BCJ,Feng:2010my}
\begin{equation}
\sum \limits_{i=2}^{n-1} k_1 \cdot (k_2+...+k_i) A_n[2,...,i,1,i+1,...,n] =0.
\end{equation}
In terms of constructing explicit color-dual solutions at tree level,
it is enough to specify the numerator of the half-ladder because any
diagram can be reduced to a half-ladder through successive
applications of the Jacobi identity.  At one-loop, every diagram can be
related to one topology as well, in this case the $n$-gon, which can be understood as a forward limit of the half-ladder.  Two-loop
calculations play an important role in our understanding of
color-kinematics duality because this is the first order where
multiple basis graphs are required, at least for a generic theory. We refer the reader to \cite{BCJreview} for more background on the topic.

@@

Here we can think of the abelian vector as a background field that
sources on-shell $A_\mu$ currents. Indeed, as we describe in
\cref{sec:CKLagrangians}, semi-abelian Yang-Mills theory is at the
heart of $YZ$-theory \cite{Cheung:2016prv}, $J$-theory
\cite{Cheung:2020djz,Cheung:2021zvb}, self-dual Yang-Mills
\cite{Monteiro2011pc}, and Chern-Simons theory \cite{Ben-Shahar:2021zww}. \jm{I mentioned CS.} \nhp{Looks good}. Specifically, semi-abelian YM is simply a
clever reinterpretation of the Lagrangian obtained by integrating an
auxiliary field into the $J$-theory equations of motion.  The Feynman
rule for the cubic vertex is
\begin{equation}\label{semiYMvert}
  \cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3)
  -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3)
\end{equation}
where the incoming arrow is the abelianized gauge field, and the outgoing arrows the on-shell non-abelian vectors. The propagator is simply
\be\label{semiYMprop}
\propFrule{$A^\nu$}{$\bar{A}^\mu$}{$k \rightarrow $}{fermion2} = \frac{i}{k^2} \eta^{\mu\nu}.
\ee
We can use the Ward identity of
\cref{eq:GIrelA} to check that the amplitudes are indeed gauge
invariant when the \emph{abelian} polarizations are taken to be longitudinal,
$\bar{\varepsilon} \rightarrow k$.
Furthermore, a simple calculation shows that the four-point correlation function of this theory satisfies the Jacobi identity of \cref{jacID} \emph{off-shell}. This ensures that color-kinematics duality holds to all multiplicity and loop order.
We also note that semi-YM does not have any ghosts from non-abelian gauge symmetry that could spoil color-kinematics at loop level. Due to the Feynman rules of this theory, the amplitudes are non-vanishing only at tree-level and one-loop. We demonstrate the implications of this property in \cref{2LoopObstruction}.

As noted above, we can select out the manifestly cubic
sector of semi-abelian Yang-Mills from the full theory of \cref{eq:ym-lag} by selecting only $(\varepsilon^+\varepsilon^+)$ in light-cone gauge (i.e., SDYM) and one-minus at tree-level or by plugging in the on-shell states of $J$-theory \cite{Cheung:2020djz,Cheung:2021zvb} and
$YZ$-theory \cite{Cheung:2016prv}. Indeed, semi-YM is just the following sum over building blocks
in the expansion of Yang-Mills given in Ref.~\cite{Pavao:2022kog},
\begin{equation}
A(A_1,...,\bar{A}_i,...,A_n)_{\text{tree}} = \sum_{i\neq j} (\varepsilon_i \varepsilon_j) \Delta^{(ij)}_{\text{tree}} \sim \sum_{i\neq j} (\varepsilon_i \varepsilon_j)  \sum_a \,[(\varepsilon k)^{n-2} (kk)^{3-n}]_a\,,
\end{equation}
and similarly so at one-loop,
\begin{equation}
A(A_1,...,A_n)_{\text{1-loop}} = \Delta^{(\varnothing)}_{\text{1-loop}} \sim \sum_a\,[ (\varepsilon k)^{n} (kk)^{-n}]_a\,,
\end{equation}
where $[\,\cdots]_a$ are terms with appropriate powers of $(\varepsilon k)$ and $(kk)$. As we describe in detail in \cref{sec:CKLagrangians}, to recover NLSM at one-loop we need to extract the $D$-dependent part
that corresponds to an internal $\bar{Y}Y$-loop from extra-dimensional,
scalars
\begin{eBox}
\begin{equation}
A^{\text{NLSM}}_{\text{1-loop}} \equiv \partial_D \Delta^{(\varnothing)}_{\text{1-loop}}\big|_{\varepsilon\rightarrow k}\, .
\end{equation}
\end{eBox}
Why must we take the derivative with respect to $D$? After all,
$J$-theory and $YZ$-theory both have well defined propagators for
internal $\bar{J}J$ and $\bar{Z}Z$ propagators. However, as we discuss
in \cref{sec:CKLagrangians}, all the $J$ and $Z$ states must be on-shell in order to
produce NLSM amplitudes. Thus, the unitarity cuts of $J$ theory at
one-loop will not produce NLSM amplitudes. However, the internal
$YY$-loop is a valid forward limit for producing NLSM amplitudes,
since as constructed $YZ$-theory matches to NLSM for off-shell
$Y$-particles. We now provide explicit expressions for these one-loop amplitudes in the next section.

@@

loop level is for self-dual Yang-Mills (SDYM). As we have done throughout the text, we will set all coupling constants to unity. At tree-level, the
off-shell cubic vertices can be written in terms of light cone
coordinates \cite{Monteiro2011pc}

@@

Before proceeding, we note that the above definition does give rise to ``pathological" bubble-on-external-leg (BEL) diagrams, discussed previously in \cref{sec:bootstrap}. However, one can show that these diagrams integrate to zero for spacetime dimension, $D>2$, and thus can be disregarded as unphysical. For the interested reader, in \cref{BELreg} we provide a detailed overview of this dimensional regularization of the relevant BEL diagram.
\subsection{Two-loop obstruction}\label{2LoopObstruction}
At two-loop, introducing terms that conspire with four-point contacts
is unavoidable. At one-loop, we were able to avoid internal
contractions of $\bar{A}_\mu A^\mu$ by selecting appropriate external
states. However, at two-loop when choosing all external $A_\mu$
states, the amplitude vanishes in semi-YM theory
\begin{equation}
\mathcal{A}^{\text{semi-YM}}_{\text{2-loop}}(A_\mu,A_\mu,A_\mu,A_\mu)=0,
\end{equation}
that is, the theory is one-loop exact.
In order to produce non-vanishing interactions, we would need to
reintroduce $D$-dimensional vertices from the full Yang-Mills Lagrangian of \cref{eq:ym-lag} that we dropped in our construction semi-abelian YM. In terms of cubic graphs, the additional interaction must necessarily have the opposite number of $\bar{A}_\mu$ and unbarred $A_\mu$ fields to that of \cref{semiYMvert}. \jm{Should we say quartic vertices with the wrong weight or something like that?  We don't have $G_L$ and $G_R$ anymore.} \nhp{Good catch -- I changed the phrasing} Reintroducing these oppositely oriented vertices allows for new internal
contractions of $\bar{A}_\mu A^\mu$,
\begin{equation}\label{2loopTension}
\mathcal{A}^{\text{YM}}_{\text{2-loop}}(A_\mu,A_\mu,A_\mu,A_\mu)=\doubleBoxObstructionA +\doubleBoxObstructionB+\cdots 
\end{equation}
where we used white dots to indicate interaction vertices of
weight\footnote{We choose the convention that the abelianized gauge
  fields carry weight $\mathcal{W}[\bar{A}] = -1$, and non-abelian
  vectors are $\mathcal{W}[\bar{A}] = +1$.}
$\mathcal{W}[\bar{A}\bar{A} A] = -1$, rather than the isolated semi-YM
vertex which always carries weight $\mathcal{W}[\bar{A}A A] =
+1$. This immediately runs into the difficulty of introducing the
four-point contact needed for color-kinematics to be satisfied on all
internal edges. Thus, by a simple weight counting argument, one can see that
including this wrong sign interactions is unavoidable at two-loop and
higher.

@@

In addition to the tree-level construction, we found that kinetic mixing
between the two gauge fields permits globally color-dual one-loop
integrands. After plugging in suitable on-shell states, the integrands
produced by semi-abelian YM correspond to those of SDYM in \cref{SDYMnGon} 
and NLSM in \cref{NLSMnGon}, along with Chern-Simons theory, which we discuss in \cref{sec:CSTheory}.  \jm{Mention CS?  Maybe something like `SDYM in \cref{SDYMnGon} 
and NLSM in \cref{NLSMnGon}, where Chern-Simons theory is discussed in \cref{sec:CSTheory}.'?} \nhp{Done.} Due to the
kinetic mixing that stems from the broken Bose symmetry, \cref{2loopTension} shows why it is
impossible to construct a two-loop integrand for semi-abelian YM theory. In \cref{2loopBoot}, we overcame this obstruction by successfully constructing two-loop NLSM numerators with an ansatz-based bootstrap. However, we found that the same approach fails for two-loop pure Yang-Mills, and pin-pointed the failure to a particular unitarity cut captured by \cref{chefsKiss}. Thus, obtaining a globally color-dual
integrand beyond one-loop then requires sacrificing more than just Bose
symmetry, and in this paper we have argued that off-shell locality\footnote{An example of this can already be found in the literature, where locality is relaxed in Ref.~\cite{Mogull:2015adi} in
  order to obtain color-dual two-loop integrands with on-shell 4D states.} must
be abandoned for color-dual constructions of multi-loop Yang-Mills. To expand the available function space, we propose that one should consider including rational functions of kinematics, rather than just polynomials.

>>>>

@@

@@

Despite the tremendous success in leveraging this duality to
compute gauge and gravity observables to high orders in perturbation
theory \cite{FiveLoopN4, GeneralizedDoubleCopyFiveLoops, Bern:2018jmv}, color-kinematics remains a conjecture at loop
level. Indeed, there are many examples in the literature where
identifying color-dual representations beyond one-loop has posed a
formidable challenge
\cite{Mogull:2015adi, Johansson:2017bfl, KalinN2TwoLoop, Bern:2015ooa}. In this work, we use
the nonlinear sigma model (NLSM) and Yang-Mills, two theories proven
to permit color-dual representations at tree-level
\cite{Feng:2010my,Cachazo:2014xea}, as case studies to advance our
understanding of the kinematic algebra at the multi-loop level.

We begin with an overview of color-kinematics duality in
\cref{background}, and define the notion of ``globally'' color-dual
integrands in \cref{sec:bootstrap}.  We then construct a manifestly
color-dual theory in \cref{sec:cubic}, which generates $D$-dimensional
color-dual $n$-point numerators at both tree-level and one-loop. When
plugging in appropriate on-shell states, we find that these numerators
underpin color-dual representations of self-dual Yang-Mills, NLSM and
Chern-Simons theory through one-loop. The construction of this theory,
which we dub \textit{semi-abelian Yang-Mills}, relies on isolating the
cubic sector of pure Yang-Mills amplitudes from the four-point
contacts that are needed to fully realize non-abelian gauge
symmetry. Despite the potency of this theory through one-loop, the
construction runs into an obstruction at two-loop, which we describe
in \cref{2LoopObstruction}.

@@

The color factor
$C_\Gamma$ is the string of structure constants $f^{abc}$ associated
with the graph $\Gamma$,\footnote{This paper is focused on theories
  with particles that transform in the adjoint.  However, the double
  copy can be applied to a far more diverse set of theories with
  particles that transform in the fundamental or more diverse
  representations, see for example \cite{Johansson:2017srf,
    Johansson2014zca, Johansson:2015oia, Johansson:2019dnu, Carrasco:2020ywq,Carrasco:2022jxn,
    Carrasco:2023vjg}.}  the denominator $d_\Gamma$
is the product of propagators associated with the graph $\Gamma$, and
the numerator $N_\Gamma$ is all of the remaining kinematic factors
including dot products of momenta and polarization vectors. In this 
form the color factors $C_\Gamma$ are not independent and are related to
one another via a set of Jacobi identities that look schematically
like
\begin{equation}
C_i+C_j+C_k=0,
\end{equation}
where $i$, $j$, and $k$ are three graphs.  Example Jacobi relations
are presented in detail in \cref{sec:bootstrap}. The duality between color and kinematics is the statement that that there exists a way to rearrange factors between
the various kinematic numerators $N_\Gamma$ such that the numerators
obey the same Jacobi identities as the color factors
\begin{equation}
N_i +N_j +N_k=0.
\end{equation}
If color-kinematics duality can be achieved at the off-shell level,
then the kinematic numerators should satisfy an algebra just like the
color factors.  In theories where the kinematic algebra is known, it
is often the algebra of volume preserving diffeomorphisms
\cite{Monteiro2011pc, Ben-Shahar:2021zww, Cheung:2020djz,
  Cheung:2021zvb, Cheung:2022mix}.
When the kinematic algebra is not known explicitly, the kinematic numerators are typically constructed from an ansatz.
As an example, the result of such a calculation for four-point NLSM scattering at tree level produces the numerators
\begin{align}\label{pionNums}
&N^{\text{NLSM}}_s = s(t-u)\\
&N^{\text{NLSM}}_t = t(u-s)\\
&N^{\text{NLSM}}_u = u(s-t)
\end{align}
where the coupling has been normalized away.  These numerators sum to
zero by explicit calculation.  Note that the $s$-channel numerator
contains an explicit factor of $s$ that cancels with the propagator
and the same goes for the $t$- and $u$-channels.  The net result is
that the amplitude is a local function, just as one would expect for
the pion four-point tree amplitude.

Once color-dual numerators $N_\Gamma$ have been found, they can be
used to replace the color factors ($C_\Gamma \to N_\Gamma$) in a
\emph{different} amplitude to produce the corresponding amplitude in
a new theory,
\be\label{DCconstruction}
\tilde{\mathcal{A}}_n^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int
\frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{C_\Gamma
  \tilde{N}_\Gamma}{d_\Gamma} ~~ \xrightarrow{(C_\Gamma \to N_\Gamma)}
~~ \mathcal{M}_n^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int
\frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{N_\Gamma
  \tilde{N}_\Gamma}{d_\Gamma}.
\end{equation}
This is known as the double-copy construction
\cite{BCJ,Bern:2010ue,BCJreview}.  Importantly, the $\tilde{N}_\Gamma$
do not need to respect color-kinematics duality.  The prototypical
example of the double-copy is that YM double copied with YM results in
gravity where the gauge invariance of each separate gluon produces the
diffeomorphism invariance of the graviton.  The double copy can be
proven at tree level using the Kawai-Lewellen-Tye (KLT) relations but
there is significant evidence that the double copy persists to all
loop orders \cite{FiveLoopN4, GeneralizedDoubleCopyFiveLoops,
  Bern:2018jmv, KLT, KiermaierTalk}.  In addition to pointing at
undiscovered structure hidden in the Lagrangian for gravity, the
double copy is immensely useful at a utilitarian level since it is
much more efficient at producing $\mathcal{M}_n^L$ than traditional
methods like Feynman rules.

Many of the one-loop results presented in \cref{sec:cubic} are informed
by the tree-level double-copy so we will review the tree case here.
The color structure of tree-level amplitudes is particularly simple.
A fully color dressed tree amplitude $\mathcal{A}_n$ can be decomposed
in terms of the trace basis
\begin{equation}
  \mathcal{A}_n = \sum \limits_{\sigma\in S_{n-1}} \text{Tr}(T^{a_{\sigma_1}} T^{a_{\sigma_2}}...T^{a_{\sigma_{n-1}}} T^{a_n}) A_n[\sigma_1, \sigma_2,...\sigma_{n-1}, n],
  \label{eq:trace-basis}
\end{equation}
where $T^a$ are generators of the gauge group and, because of the
cyclicity of the trace, one of the legs is held fixed so that the sum
runs over the $(n-1)!$ permutations of the remaining external legs.
The coefficients of the color factors are the color-ordered partial
amplitudes $A_n[...]$.  Any tree-level color factor can be converted
into a linear combination of Del Duca-Dixon-Maltoni (DDM) half-ladder
color factors
\begin{equation}
\text{DDM}[a_1, a_2,... a_n] \equiv f^{a_1 a_2 b_1} f^{b_1 a_3 b_2} f^{b_2 a_4 b_3}...f^{b_{n-3} a_{n-1} a_n}
\end{equation}
associated with the graph
\begin{equation}
\halfLadder
\end{equation}
by repeated application of the Jacobi identity \cite{DixonMaltoni}.
The fully color-dressed tree amplitude can then be re-expressed as
\begin{equation}
  \mathcal{A}_n = \sum \limits_{\sigma\in S_{n-2}} \text{DDM}[a_1, a_{\sigma_2}, a_{\sigma_3}, a_{\sigma_{n-1}}, a_n]  A_n[1, \sigma_2,...\sigma_{n-1}, n]
  \label{eq:ddm-basis}
\end{equation}
with legs $1$ and $n$ fixed.  Since the sum only contains $(n-2)!$
terms, there must be additional relations -- the
Kleiss-Kuijf (KK) relations \cite{Kleiss:1988ne} -- amongst the
partial amplitudes appearing in the right-hand side of
\cref{eq:trace-basis}.  The KK relations are generic to any theory
with purely adjoint particles.  Color-kinematics duality further
implies that the KK basis is overcomplete and can be reduced to a
basis of $(n-3)!$ amplitudes via the fundamental Bern-Carrasco-Johansson (BCJ) identities
\cite{BCJ,Feng:2010my}
\begin{equation}
\sum \limits_{i=2}^{n-1} k_1 \cdot (k_2+...+k_i) A_n[2,...,i,1,i+1,...,n] =0.
\end{equation}
In terms of constructing explicit color-dual solutions at tree level,
it is enough to specify the numerator of the half-ladder because any
diagram can be reduced to a half-ladder through successive
applications of the Jacobi identity.  At one-loop, every diagram can
be related to one topology as well, in this case the $n$-gon, which
can be understood as a forward limit of the half-ladder.  Two-loop
calculations play an important role in our understanding of
color-kinematics duality because this is the first order where
multiple basis graphs are required, at least for a generic theory. We
refer the reader to Ref.~\cite{BCJreview} for more background on the
topic.

@@

Here we can think of the abelian vector as a background field that
sources on-shell $A_\mu$ currents. Indeed, as we describe in
\cref{sec:CKLagrangians}, semi-abelian Yang-Mills theory is at the
heart of $YZ$-theory \cite{Cheung:2016prv}, $J$-theory
\cite{Cheung:2020djz,Cheung:2021zvb}, self-dual Yang-Mills
\cite{Monteiro2011pc}, and Chern-Simons theory
\cite{Ben-Shahar:2021zww}. Specifically, semi-abelian YM is simply a
clever reinterpretation of the Lagrangian obtained by integrating an
auxiliary field into the $J$-theory equations of motion.  The Feynman
rule for the cubic vertex is
\begin{equation}\label{semiYMvert}
  \cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3)
  -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3)
\end{equation}
where the incoming arrow is the abelianized gauge field, and the outgoing arrows the on-shell non-abelian vectors. The propagator is simply
\be\label{semiYMprop}
\propFrule{$A^\nu$}{$\bar{A}^\mu$}{$k \rightarrow $}{fermion2} = \frac{i}{k^2} \eta^{\mu\nu}.
\ee
We can use the Ward identity of
\cref{eq:GIrelA} to check that the amplitudes are indeed gauge
invariant when the \emph{abelian} polarizations are taken to be longitudinal,
$\bar{\varepsilon} \rightarrow k$.
Furthermore, a simple calculation shows that the four-point correlation function of this theory satisfies the Jacobi identity of \cref{jacID} \emph{off-shell}. This ensures that color-kinematics duality holds to all multiplicity and loop order.
We also note that semi-YM does not have any ghosts from non-abelian gauge symmetry that could spoil color-kinematics at loop level. Due to the Feynman rules of this theory, the amplitudes are non-vanishing only at tree-level and one-loop. We demonstrate the implications of this property in \cref{2LoopObstruction}.

As noted above, we can select out the manifestly cubic
sector of semi-abelian Yang-Mills from the full theory of \cref{eq:ym-lag} by selecting only $(\varepsilon^+\varepsilon^+)$ in light-cone gauge (i.e., SDYM) and one-minus at tree-level or by plugging in the on-shell states of $J$-theory \cite{Cheung:2020djz,Cheung:2021zvb} and
$YZ$-theory \cite{Cheung:2016prv}. Indeed, semi-YM is just the following sum over building blocks
in the expansion of Yang-Mills given in Ref.~\cite{Pavao:2022kog},
\begin{equation}
A(A_1,...,\bar{A}_i,...,A_n)_{\text{tree}} = \sum_{i\neq j} (\varepsilon_i \varepsilon_j) \Delta^{(ij)}_{\text{tree}} \sim \sum_{i\neq j} (\varepsilon_i \varepsilon_j)  \sum_a \,[(\varepsilon k)^{n-2} (kk)^{3-n}]_a\,,
\end{equation}
and similarly so at one-loop,
\begin{equation}
A(A_1,...,A_n)_{\text{1-loop}} = \Delta^{(\varnothing)}_{\text{1-loop}} \sim \sum_a\,[ (\varepsilon k)^{n} (kk)^{-n}]_a\,,
\end{equation}
where $[\,\cdots]_a$ are terms with appropriate powers of
$(\varepsilon k)$ and $(kk)$. As we describe in detail in
\cref{sec:CKLagrangians}, to recover NLSM at one-loop we need to
extract the $D$-dependent part that corresponds to an internal
$\bar{Y}Y$-loop from extra-dimensional scalars
\begin{eBox}
\begin{equation}
A^{\text{NLSM}}_{\text{1-loop}} \equiv \partial_D \Delta^{(\varnothing)}_{\text{1-loop}}\big|_{\varepsilon\rightarrow k}\, .
\end{equation}
\end{eBox}
Why must we take the derivative with respect to $D$? After all,
$J$-theory and $YZ$-theory both have well defined propagators for
internal $\bar{J}J$ and $\bar{Z}Z$ propagators. However, as we discuss
in \cref{sec:CKLagrangians}, all the $J$ and $Z$ states must be on-shell in order to
produce NLSM amplitudes. Thus, the unitarity cuts of $J$ theory at
one-loop will not produce NLSM amplitudes. However, the internal
$YY$-loop is a valid forward limit for producing NLSM amplitudes,
since as constructed $YZ$-theory matches to NLSM for off-shell
$Y$-particles. We now provide explicit expressions for these one-loop amplitudes in the next section.

@@

loop level is for self-dual Yang-Mills (SDYM). As we have done
throughout the text, we will set all coupling constants to unity. At
tree-level, the off-shell cubic vertices can be written in terms of
light cone coordinates \cite{Monteiro2011pc}

@@

Before proceeding, we note that the above definition does give rise to ``pathological" bubble-on-external-leg (BEL) diagrams, discussed previously in \cref{sec:bootstrap}. However, one can show that these diagrams integrate to zero for spacetime dimension, $D>2$, and thus can be disregarded as unphysical. For the interested reader, in \cref{BELreg} we provide a detailed overview of this dimensional regularization of the relevant BEL diagram.
\subsection{Two-loop obstruction}\label{2LoopObstruction}
At two-loop, introducing terms that conspire with four-point contacts
is unavoidable. At one-loop, we were able to avoid internal
contractions of $\bar{A}_\mu A^\mu$ by selecting appropriate external
states. However, at two-loop when choosing all external $A_\mu$
states, the amplitude vanishes in semi-YM theory
\begin{equation}
\mathcal{A}^{\text{semi-YM}}_{\text{2-loop}}(A_\mu,A_\mu,A_\mu,A_\mu)=0,
\end{equation}
that is, the theory is one-loop exact.  In order to produce
non-vanishing interactions, we would need to reintroduce
$D$-dimensional vertices from the full Yang-Mills Lagrangian of
\cref{eq:ym-lag} that we dropped in our construction semi-abelian
YM. In terms of cubic graphs, the additional interaction must
necessarily have the opposite number of $\bar{A}_\mu$ and unbarred
$A_\mu$ fields to that of \cref{semiYMvert}.  Reintroducing these
oppositely oriented vertices allows for new internal contractions of
$\bar{A}_\mu A^\mu$,
\begin{equation}\label{2loopTension}
\mathcal{A}^{\text{YM}}_{\text{2-loop}}(A_\mu,A_\mu,A_\mu,A_\mu)=\doubleBoxObstructionA +\doubleBoxObstructionB+\cdots 
\end{equation}
where we used white dots to indicate interaction vertices of
weight\footnote{We choose the convention that the abelianized gauge
  fields carry weight $\mathcal{W}[\bar{A}] = -1$, and non-abelian
  vectors are $\mathcal{W}[\bar{A}] = +1$.}
$\mathcal{W}[\bar{A}\bar{A} A] = -1$, rather than the isolated semi-YM
vertex which always carries weight $\mathcal{W}[\bar{A}A A] =
+1$. This immediately runs into the difficulty of introducing the
four-point contact needed for color-kinematics to be satisfied on all
internal edges. Therefore, by a simple weight counting argument, one can see that
including this wrong sign interactions is unavoidable at two-loop and
higher.

@@

In addition to the tree-level construction, we found that kinetic
mixing between the two gauge fields permits globally color-dual
one-loop integrands. After plugging in suitable on-shell states, the
integrands produced by semi-abelian YM correspond to those of SDYM in
\cref{SDYMnGon} and NLSM in \cref{NLSMnGon}, along with Chern-Simons
theory, which we discuss in \cref{sec:CSTheory}.  Due to the kinetic
mixing that stems from the broken Bose symmetry, \cref{2loopTension}
shows why it is impossible to construct a two-loop integrand for
semi-abelian YM theory. In \cref{2loopBoot}, we overcame this
obstruction by successfully constructing two-loop NLSM numerators with
an ansatz-based bootstrap. However, we found that the same approach
fails for two-loop pure Yang-Mills, and pin-pointed the failure to a
particular unitarity cut captured by \cref{chefsKiss}. Thus, obtaining
a globally color-dual integrand beyond one-loop then requires
sacrificing more than just Bose symmetry, and in this paper we have
argued that off-shell locality\footnote{An example of this can already
  be found in the literature, where locality is relaxed in
  Ref.~\cite{Mogull:2015adi} in order to obtain color-dual two-loop
  integrands with on-shell 4D states.} must be abandoned for
color-dual constructions of multi-loop Yang-Mills. To expand the
available function space, we propose that one should consider
including rational functions of kinematics, rather than just
polynomials.
